A Dependently Typed Multi-Stage Calculus\thanks{Supported by organization x.}

### 2019-05-27 main.tex
Filling gaps.
Finish section 4.

Signatures are sequences of pairs of a constant and its type or a type variable and its kind.
Because we adopted a \LLF-like system, constants and type variables for base types are given in the signature $\Sigma$.

@@

In this section, we define the type system of \LMD.
Especially, we discuss on the type equality of \LMD in detail
because it is the main contribution of this paper.
\LMD is developed on the basis of \LTP but there are no type equivalence rules in \LTP.

@@

@@

@@

@@

\begin{theorem}[Confluence(Church-Rosser Property)]
	Define $M \longrightarrow N$ as $M \longrightarrow_{\beta} N$ or $M\longrightarrow_\blacklozenge N$ or  $M \longrightarrow_{\Lambda} N$.\\
	For any term $M$, if $M \longrightarrow^* N$ and $M \longrightarrow^* L$,
	there exists $O$ that satisfies $N \longrightarrow^* O$ and $L \longrightarrow^* O$.
\end{theorem}

\begin{theorem}[Progress]
	If $x:\tau @\epsilon \notin \G$ and $\G \V M : \tau  @ A$ then $ M \in V^A $ or $\exists M'$ such that $M \longrightarrow M'$.
\end{theorem}

\begin{theorem}[Unique Decomposition]
	If $x:\tau@\epsilon \notin \G$ and $\G \V M : \tau @ A$ then 1 or 2 is true.
	\begin{enumerate}
		\item $ M \in V^A$
		\item $\exists ! B, E^A_B, R^B$ such that ($B = \epsilon$ or $B = \beta$) and $M = E^A_B[R^B]$.
	\end{enumerate}
\end{theorem}

>>>>

Signatures are sequences of pairs of a constant and its type or a type variable and its kind.
Because we adopted a \LLF-like system, constants and type variables for base types are given in the signature $\Sigma$.
For example, when we use integers in \LMD, $\Sigma = \I :: *, 1:\I, 2:\I, \cdots$.

@@

In this section, we define the type system of \LMD.
\LMD\ is little complicated because it contains dependent types.
It contains typing, kinding, well-formed kinding, term equality, type equality, and kind equality.

Especially, we discuss on the type equality of \LMD in detail
because it is the main contribution of this paper.
\LMD is developed on the basis of \LTP but there are no type equivalence rules in \LTP.

@@

This kind of type equality is very useful when we combine $\text{mulmat}\ 3\ 5$ with another code.
We cannot remove $\%_\alpha$ in the type without \QPercent and it makes very difficult to combine
because types of codes don't contains $\%_\alpha$ generally.

@@

\begin{theorem}[Inversion Lemma for $\Pi$ type]
	If $\G \V (\lambda x:\sigma.M) : (\Pi x:\sigma'.\tau)@A$ then
	\begin{enumerate}
		\item $\G \V \sigma \E \sigma'@A$
		\item $\G ,x:\sigma@A\V M:\tau@A$
	\end{enumerate}
	\item If $\G \V \rho \E (\Pi x:\sigma.\tau) : K @A$ then $\exists \sigma', \tau', K, J$ such that
	\begin{enumerate}
		\item $\rho = \Pi x:\sigma'.\tau'$
		\item $\G \V \sigma \E \sigma' : K @A$
		\item $\G, x:\sigma@A\V \tau \E \tau' : J @A$
	\end{enumerate}
	\item If $\G \V (\Pi x:\sigma.\tau) \E \rho : K @A$ then $\exists \sigma', \tau', K, J$ such that
	\begin{enumerate}
		\item $\rho = \Pi x:\sigma'.\tau'$
		\item $\G \V \sigma \E \sigma' : K @A$
		\item $\G, x:\sigma@A\V \tau \E \tau' : J @A$
	\end{enumerate}
\end{theorem}

\begin{theorem}[Inversion Lemma for $\TW$ type]
	\begin{item}
	      \item If $\G \V \TB_\alpha M : \TW_\alpha \tau@A$ then $\G \V M : \tau @A$.
	      \item If $\G \V \rho \E  \TW_\alpha \tau : K @A$ then $\exists \tau', K, J$ such that
	      \begin{enumerate}
		      \item $\rho = \TW_\alpha \tau'$
		      \item $\G @A\V \tau \E \tau' : K @A$
	      \end{enumerate}
	      \item If $\G \V \TW_\alpha \tau \E \rho : K @A$ then $\exists \tau', K, J$ such that
	      \begin{enumerate}
		      \item $\rho = \TW_\alpha \tau'$
		      \item $\G @A\V \tau \E \tau' : K @A$
	      \end{enumerate}
	\end{item}
\end{theorem}

\begin{theorem}[Inversion Lemma for $\Lambda$ type]
	\begin{item}
	      \item If $\G \V \Lambda\alpha.M : \forall\alpha.\tau@A$ then $\G \V M : \tau @A$ and $\alpha \notin \FTV(\G) \cup \FV(A)$.
	      \item If $\G \V \rho \E \forall\alpha.\tau : K @A$ then $\exists \tau', K$ such that
	      \begin{enumerate}
		      \item $\rho = \forall\alpha.\tau'$
		      \item $\G \V \tau \E \tau' : K @A$
	      \end{enumerate}
	      \item If $\G \V \forall\alpha.\tau \E \rho : K @A$ then $\exists \tau', K$ such that
	      \begin{enumerate}
		      \item $\rho = \forall\alpha.\tau'$
		      \item $\G \V \tau \E \tau' : K @A$
	      \end{enumerate}
	\end{item}
\end{theorem}

@@

\begin{proof}
	First, there are three cases for $M \longrightarrow M'$.
	They are $M \longrightarrow_\beta M'$, $M \longrightarrow_\Lambda M'$, and $M \longrightarrow_\blacklozenge M'$.
	For each case, we can use straightforward induction on derivations.
	Difficult cases are \TApp, \TTBL, and \TIns.
	We need Inversion Lemmas for them.
\end{proof}

@@

\begin{proof}
	In order prove this theorem, we define a translation $\natural$ from \LMD\ to Simply Typed Lambda Calculus.
	Second, we prove $\natural$ preserves typing and $\beta$-reductions.
	Then, we can prove Strong Normalization of \LMD\ from Strong Normalization of Simply Typed Lambda Calculus.
\end{proof}

\begin{theorem}[Confluence]
	For any term $M$, if $M \longrightarrow^* M'$ and $M \longrightarrow^* M''$ then
	there exists $M'''$ that satisfies $M' \longrightarrow^* M'''$ and $M'' \longrightarrow^* M'''$.
\end{theorem}

\begin{proof}
	Because we proved Strong Normalization of \LMD, 
	we can use Newman's lemma to prove Confluence of \LMD.
	Then, what we must show is Weak Church-Rosser Property now.
	When we consider two different redexes in a \LMD term, they can only be disjoint, or one is a part of the other.
	In short, they are never overlapped each other.
	So, we can reduce one of them after we reduce another.
\end{proof}

\begin{theorem}[Progress]
	If $x:\tau @\epsilon \notin \G$ and $\G \V M : \tau  @ A$ then
	$ M \in V^A $ or $M'$ exists such that $M \longrightarrow M'$.
\end{theorem}

\begin{proof}
	We can prove by straightforward induction on derivations.
	Difficult cases are \TApp, \TTBL, and \TIns.
	We need Inversion Lemmas for them.
\end{proof}

\begin{theorem}[Unique Decomposition]
	If $x:\tau@\epsilon \notin \G$ and $\G \V M : \tau @ A$ then 1 or 2 is true.
	\begin{enumerate}
		\item $ M \in V^A$
		\item There exist an unique tuple of $B, E^A_B, R^B$ such that $M = E^A_B[R^B]$ ($B = \epsilon$ or $B = \beta$) .
	\end{enumerate}
\end{theorem}

\begin{proof}
	We can prove by straightforward induction on derivations.
	Difficult cases are \TApp, \TTBL, and \TIns.
	We need Inversion Lemmas for them.
\end{proof}
